---
title: "Applied Statistics: Lab 2"
author: "Yangzom Dolma"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

# Task 1

```{r}
getwd()
```

# Task 2

```{r}
epagas = read.csv("EPAGAS.csv")
head(epagas)
```

# Task 3

Make the object mpg, the number of miles per gallon vector.

```{r}
mpg <- epagas$MPG
head(mpg)
```

z-score of MPG: 
```{r}
z = (mpg - mean(mpg))/sd(mpg)
head(z)
```

Verifying that mean(z) = 0 and that standard deviation of z is 1.

```{r}
round(mean(z))
sd(z)
```

Possible outliers for the values of mpg: 

```{r}
poutliers = mpg[abs(z) >= 2 & abs(z) <= 3]
poutliers
```

The values of mpg that are outliers: 

```{r}
outlier = mpg[abs(z) > 3]
outlier
```

Dotplot using lattice package: 

```{r}
library(lattice)
mpgplot = ifelse(abs(z) > 3, "red", 
                 ifelse(abs(z) >=2 & abs(z) <= 3, "blue", "black"))
dotplot(mpg, col = mpgplot, main = "Dotplot of MPG")
```

# Task 4

Boxplot of MPG data: 

```{r}
boxplot(mpg, 
        main = "Boxplot of MPG",
        col = "black",
        xlab = "MPG",
        horizontal = TRUE,
        notch = TRUE)
```

### Chebyshev's Theorem

Using Chebyshev's Theorem: Proportion of MPG data within 2 standard deviations of the mean. 

```{r}
k = 2
ch = (1 - 1/k^2)
ch
```

Using R: Proportion of MPG data within 2 standard deviations of the mean.

```{r}
twoSD = mpg[abs(z) < 2]
exact = length(twoSD)/length(mpg)
exact
```

Does Chebyshev agree with the data?

Ans: Yes, Chebyshev agree with the data. Chebyshev indicates that at least 75% of the MPG data will be within 2 standard deviations of the mean. The exact proportion of data within 2 standard deviations of the mean is found to be 96% which is more than 75%. 

### Empirical Rule

According to the empirical rule, approximately 95% of the data should be within 2 standard deviations of the mean of the data. 

In this case, we found that 96% of the data is within 2 standard deviations of the mean, which correspond approximately close to 95%. 

Density Distribution Plot of MPG data: 

```{r}
plot(density(mpg), main = "Empirical Distribution of MPD Data", col = "blue")
abline (v = mean(mpg), col = "green")
```

The density distribution plot of MPG data is mound-shaped and has symmetric distribution. Therefore,in this case the empirical rule is valid. 

