---
title: "Math 4753 Lab1: Introduction to R"
author: "Yangzom Dolma"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

# Task 1

The working directory: 

```{r}
getwd() 
```

# Task 2

First six lines of data. 

```{r}
ddt = read.csv("DDT.csv") 
head(ddt) 
```
# Task 3

a. The qualitative variables in ddt data set are:

Ans: RIVER and SPECIES. 

b. The quantitative variables in the ddt data set are: 

Ans: MILE, LENGTH, WEIGHT and DDT. 

c. Number of species in the ddt data set.  

Ans. 3 Species.

```{r}
df<-with(ddt, SPECIES) 
df_unique<-unique(df) 
length(df_unique) 
```

d. Subset the ddt data set so that we see only the observations with SPECIES,  LMBASS and WEIGHT bigger than 800gms. 
```{r}
ddt_subset<-with(ddt, ddt[WEIGHT>800 & SPECIES=="LMBASS",]) 
head (ddt_subset) 
```

e. Subset the ddt data set so that we see only those observations where the RIVER is SCM   and DDT bigger than 4.0

```{r}
ddt_river <- with(ddt, ddt[DDT > 4.0 & RIVER == "SCM",]) 
head (ddt_river) 
```

# Clicker Questions

Q1. The mean length of fish in the ddt data frame. 

Ans: 42.81 length unit. 

```{r}
z<-mean(ddt$LENGTH) 
z 
```


Q2. What is the standard deviation of the weight of fish in the ddt data frame?

Ans: 376.5461

```{r}
x <- sd(ddt$WEIGHT) 
x  
```


Q3. Is this the plot of LENGTH vs WEIGHT from the ddt data frame? 

Ans:  1 = Yes

Q4. If v=1:20 what is the last value of v/20?

Ans: 1.00

```{r}
v = 1:20  
v/20 
```


# Task 4: Table and Barplot

a. Table of the rivers.

```{r}
RiverTable = table(ddt$RIVER) 
print(RiverTable) 
```

b. Color coded barplot of the rivers.

```{r}
barplot(table(ddt$RIVER), col = 1:4) 
```


c. Table of the RIVERS crossed with the SPECIES of fish.

```{r}
tableRS <- with(ddt, table(SPECIES, RIVER)) 
print(tableRS) 
```

d. Barplot of the RIVER crossed with the SPECIES of fish.

```{r}
barplot(tableRS, col = rainbow(3),
        main = "Barplot",
        border = "Black"
        )
```

# Task 5: Pie Chart

a. Pie chart of Fish SPECIES

```{r}
fish = ddt$SPECIES 
pie (table(fish)) 
```

b. Pie chart of RIVERS

```{r}
rivers = ddt$RIVER 
pie (table(rivers)) 
```


# Task 6: Boxplot

a. Boxplot of the variable DDT

```{r}
boxplot(ddt$DDT, 
        main = "Boxplot of DDT",
        xlab = "dbl",
        ylab = "DDT", 
        col = "Orange",
        border = "brown",
        horizontal = TRUE,
        notch = TRUE
) 
```

b. Boxplot of the variable WEIGHT

```{r}
boxplot(ddt$WEIGHT, 
        main = "Boxplot of WEIGHT",
        xlab = "weight unit",
        ylab = "WEIGHT", 
        col = "Orange",
        border = "brown",
        horizontal = TRUE,
        notch = TRUE
) 
```


c. Boxplot of the variable LENGTH

```{r}
boxplot(ddt$LENGTH, 
        main = "Boxplot of LENGTH",
        xlab = "Length unit",
        ylab = "LENGTH", 
        col = "Orange",
        border = "brown",
        horizontal = TRUE,
        notch = TRUE
) 
```


# Task 7: Coplot

a. Coplot of LENGTH Vs WEIGHT given RIVER

```{r}
coplot(LENGTH ~ WEIGHT | RIVER, 
       data=ddt, 
       col = with(ddt, ifelse (SPECIES == "CCATFISH", "Pink",
                  ifelse(SPECIES == "SMBUFFALO", "Green", "Orange")))) 
```


b. Coplot of DDT vs WEIGHT given SPECIES


```{r}
coplot(DDT ~ WEIGHT | SPECIES, data = ddt) 
```


# Task 8: ggplot: Box, Violin, Scatter

a. ggplot: Boxplot for WEIGHT vs SPECIES

```{r}
library(ggplot2)
ggplot(ddt, ggplot2::aes(x=SPECIES, y=WEIGHT)) + geom_boxplot(aes(fill=RIVER)) + ggtitle("Yangzom Dolma") 
```
b. ggplot: Violin Plot for LENGTH vs RIVER

```{r}
print(ggplot(ddt, aes(x=RIVER, y = LENGTH)) + geom_violin(aes(fill =SPECIES)) + ggtitle("Yangzom Dolma")) 
```
c. ggplot: Scatter Plot for LENGTH vs WEIGHT

```{r}
ggsc = ggplot(ddt, aes(x= WEIGHT, y = LENGTH))

ggsc = ggsc + geom_point(aes(color = SPECIES)) + ggtitle("Yangzom Dolma")

print(ggsc)
```

