---
title: "Math 4753 Lab 5"
author: "Yangzom Dolma"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

# Task 1

Working directory: 

```{r}
getwd()
```

# Task 2

## Binomial Experiment

```{r}

mybin=function(iter=100,n=10, p=0.5){ 
  
#initially filled with NA's to hold the samples
sam.mat = matrix(NA, nrow = n, ncol = iter, byrow = TRUE)

#Make a vector to hold the number of successes in each trial
succ=c()

#table of success
for( i in 1:iter){
  
#Fill each column with a new sample
sam.mat[,i]=sample(c(1,0),n,replace=TRUE, prob=c(p,1-p))

#Calculate a statistic from the sample (this case it is the sum)
succ[i]=sum(sam.mat[,i])
}

#Make a table of successes
succ.tab=table(factor(succ,levels=0:n))

#Make a barplot of the proportions

iter.lab = paste0("iter = ", iter)
n.lab = paste0("n = ", n)
p.lab = paste0("p = ", p)
lab = paste(iter.lab, n.lab, p.lab, sep = "   ")
barplot(succ.tab/(iter), col=rainbow(n+1), main="Binomial simulation",sub = lab, xlab="Number of successes")
succ.tab/iter
}
```

### 100 iterations

```{r}
mybin100 = mybin(iter=100,n=10, p=0.7)
```
### 200 iterations

```{r}
mybin200 = mybin(200, 10, 0.7)
```

### 500 iterations

```{r}
mybin500 = mybin(500, 10, 0.7)
```

### 1000 iterations

```{r}
mybin1000 = mybin(1000, 10, 0.7)
```

### 10, 000 iterations and its table

```{r}
mybin(10000, 10, 0.7)
```

## Verify table values are approximate

Looking at the values after the 10,000 iterations and the values from dbinom, we can conclude that they are approximately correct. 

```{r}
dbin.tab = round(dbinom(0:10, size = 10, p = 0.7), 4)
names(dbin.tab) = 0:10
dbin.tab

```


# Task 3

## Sample without replacement

```{r}
sample(c(1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0), size = 5, replace = FALSE, prob = NULL)
```

## Sample with replacement

```{r}
sample(c(1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0), size = 5, replace = TRUE, prob = NULL)
```

## hyper function

```{r}
myhyper=function(iter=100, N=20, r = 12, n = 5){ 
  
#initially filled with NA's to hold the samples
sam.mat = matrix(NA, nrow = n, ncol = iter, byrow = TRUE)

#Make a vector to hold the number of successes in each trial
succ=c()

#table of success
for( i in 1:iter){
  
#Fill each column with a new sample
sam.mat[,i]=sample(rep(c(1,0), c(r, N-r)), n, replace = FALSE)

#Calculate a statistic from the sample (this case it is the sum)
succ[i]=sum(sam.mat[,i])
}

#Make a table of successes
succ.tab=table(factor(succ,levels=0:n))

#Make a barplot of the proportions

iter.lab = paste0("iter = ", iter)
N.lab = paste0("N = ", N)
r.lab = paste0("r = ", r)
n.lab = paste0("n = ", n)
lab = paste(iter.lab, N.lab, r.lab, n.lab, sep = "   ")
barplot(succ.tab/(iter), col=rainbow(n+1), main="Hypergeometric simulation",sub = lab, xlab="Number of successes")
succ.tab/iter
}
```


## myhyper: 100 iterations

```{r}
hg1 = myhyper(iter = 100, N =20, r = 12, n =5)
```

## myhyper: 200 iterations

```{r}
hg1 = myhyper(iter = 200, N =20, r = 12, n =5)
```

## myhyper: 500 iterations

```{r}
hg3 = myhyper(iter = 500, N =20, r = 12, n =5)
```

## myhyper: 1000 iterations

```{r}
hg4 = myhyper(iter = 1000, N =20, r = 12, n =5)
```

## 10, 000 iterations and table produced

```{r}
myhyper(iter = 10000, N =20, r = 12, n =5)
```

## Verify values are appromixate

We can verify that the values from dhyper and the values after 10,000 iterations are approximately correct.

```{r}
dhyper.tab = round (dhyper(x=0:5, m =12, n=8, k=5), 4)
names(dhyper.tab) = 0:5
dhyper.tab
```

# Task 4

## mysample() description

Looking at the code for mysample function, it takes in 3 parameters namely; n, iter, and time. The iter parameter is used to set the number of iterations of the for loop. The function creates a replaceable sample with values ranging between 1 and 10 with a sample size of n. Each iteration of the for loop creates a sample of size n by choosing a random number between 1 and 10. After the sample is created, a factor variable sf is created to count the occurrences of each number. Finally a bar plot of the run is created. The y-axis value of the Bar Plot is limited by 0.02. The table is released and the delay in between each loop iteration is given by the value of the parameter time. 

```{r}
mysample=function(n= 10, iter=5,time=0.5){
for( i in 1:iter){
#make a sample
s=sample(1:10,n,replace=TRUE)
# turn the sample into a factor
sf=factor(s,levels=1:10)

#make a barplot
barplot(table(sf)/n,beside=TRUE,col=rainbow(10), 
main=paste("Example sample()", " iteration ", i, " n= ", n,sep="") ,
ylim=c(0,0.2)
)

#release the table
Sys.sleep(time)
}
}

# mysample(n=4, iter=5)
```

## Run mysample(n=1000, iter=30,time=1)

```{r}
mysample(n=1000, iter=30,time=1)
```

## Describe what you see here.

A bar plot is created every 1 second. The total number of Bar Plot created is 30 as it is the value passed on to the iter parameter of the function. The distribution for each sample is shown in the plots and they are valued at about 0.10. The y-axis of the bar plot is is limited to 0.2.


# Task 5

## Calculations

### c(8,4)

```{r}
choose(8, 4)
```

### P(Y > 4), Y~Pois(λ = 2)

```{r}
1-ppois(q=3, lambda = 2)
```

## More Calculations

### P(Y=10),Y~NegBin(p=0.4,r=3)

```{r}
choose(10-1, 3 -1)*0.4^3*0.6^(10-3)
```

```{r}
nbp = dnbinom(x = 10-3, size = 3, prob = 0.4)
nbp
```

### 	P(Y≤8),Y~Bin(n=15,p=0.4)

```{r}
pbinom(8, 15, 0.4)
```


# Task 6

## R function by copying code from this lab. 

```{r}
Intro2R::mybin()
```


# Task 7

## Simulate a neg binomial distribution


```{r}
#P(Y=10), Y~NegBin(p=0.4,r=3).   (Book theory)
# No Failures, No of successes, prob success
dnbinom(7,3,0.4)  
mynbin=function(y,r,p){
choose(y-1,r-1)*p^r*(1-p)^(y-r)
}
mynbin(10,3,0.4)
```



